Effect of multiple reflection of the sky signal off the HERA element on the measurements of 21cm power spectrum

### 2015-11-04 DishReflectometry/Paper1_v0.tex

Since it was first proposed in \citet{shaver_et_al1999}, measuring 21~cm
emission from neutral hydrogen in our early universe has gained attention as a
powerful probe of both cosmology and astrophysics.  While the science case for
21~cm cosmology, particularly during the Epoch of Reionization, is well
established (see, e.g.,
\citealt{furlanetto_et_al2006,morales_wyithe2010,furlanetto_loeb2014,pritchard_loeb2014}),
the technical path toward measuring this signal has been more problematic.  The
weakness of this hyperfine line keeps the 21~cm signal below opacity throughout
cosmological history, but it also creates sensitivity and calibration
challenges that have yet to be fully solved.  With noise temperatures dominated
by sky noise \citep{XXX} and foregrounds four to five orders of magnitude
brighter than the signal \citep{XXX}, 21~cm reionization experiments must
furnish both collecting area and foreground suppression at levels significantly
beyond anything previously achieved in radio telescopes operating below 1 GHz.

One of the most problematic effects facing these experiments results form instrument
chromaticity.  The spectral dimension of 21~cm reionization experiments is of vital
importance; for line emission, this coordinate translates to a line-of-sight distance
that can be used to construct three-dimensional maps (and power spectra) of emission,
as well as probe the evolution of 21~cm emission over cosmological timescales.
The evolving response of a radio telescope --- either a single dish
or an interferometer --- versus spectral frequency modulates spectrally smooth foregrounds ,
contaminating spectral modes that might
otherwise be used to measure reionization \citep{XXX}.  Moreover, the chromaticity of
a telescope's response pattern scales linearly with diameter, putting the
needs of foreground suppression and signal sensitivity in direct tension with one
another.

% XXX the wedge

>>>>

Since it was first proposed in \citet{shaver_et_al1999}, measuring 21~cm
emission from neutral hydrogen in our early universe has gained attention as a
powerful probe of both cosmology and astrophysics.  While the science case for
21~cm cosmology, particularly during the Epoch of Reionization, is well
established (see, e.g.,
\citealt{furlanetto_et_al2006,morales_wyithe2010,furlanetto_loeb2014,pritchard_loeb2014}),
the technical path toward measuring this signal has been more problematic.  The
weakness of this hyperfine line keeps the 21~cm signal below opacity throughout
cosmological history, but it also creates sensitivity and calibration
challenges that have yet to be fully solved.  With noise temperatures dominated
by sky noise \citep{XXX} and foregrounds four to five orders of magnitude
brighter than the signal \citep{XXX}, 
sky-averaged 21~cm monopole experiments such as
(EDGES; \citealt{XXX}),
(BIGHORNS; \citealt{XXX}),
(SARAS; \citealt{patra_et_al2014}),
(SCIHI; \citealt{voytek_et_al2014}),
(HyPerion; \citealt{presley_et_al2015})
% XXX others?
and 21~cm reionization power spectrum experiments such as
the LOw Frequency ARray (LOFAR; \citealt{XXX}),
the Murchison Widefield Array (MWA; \citealt{XXX}),
the Giant Metre-wave Radio Telescop (GMRT; \citealt{XXX}),
the Donald C. Backer Precision Array to Probe the Epoch of Reionization (PAPER; \citealt{parsons_et_al2010}),
the Hydrogen Epoch of Reionization Array (HERA; \citealt{XXX}),
and the future Square Kilometre Array (SKA; \citealt{XXX})
must
furnish both collecting area and foreground suppression at levels significantly
beyond anything previously achieved in radio telescopes operating below 1 GHz.

One of the most problematic effects facing these experiments is instrument
chromaticity.  The spectral dimension of 21~cm reionization experiments is of vital
importance; for line emission, this coordinate translates to a line-of-sight distance
that can be used to construct three-dimensional maps (and power spectra) of emission,
as well as probe the evolution of 21~cm emission over cosmological timescales.
The evolving response of a radio telescope --- either a single dish
or an interferometer --- versus spectral frequency modulates spectrally smooth foregrounds,
contaminating spectral modes that might
otherwise be used to measure reionization \citep{XXX}.  Moreover, the chromaticity of
a telescope's response scales linearly with diameter, putting the
needs of foreground suppression and signal sensitivity in direct tension with one
another.

A major step forward for the field of 21~cm cosmology has been the development
of a mathematical description of telescope chromaticity, how it varies with
element separation (or telescope diameter), and what it implies for distinguishing
foreground emission from the cosmological 21~cm signal \citep{XXX}.  The ``wedge'', as it is
colloquially known, describes a linear relationship between the separation between elements
in an interferometric baseline and the maximum line-of-sight Fourier mode\footnote{Assuming
a flat sky, and using appropriate
cosmological scalars,
the spectral axis, $\nu$, and angle on the sky, $\vec\theta$, translate to coordinates in
a three-dimensional volume at cosmological distances.  In describing the spatial power spectrum of
emission in this volume $P(\vec k)$, we use the three-dimensional wave vector 
$\vec k\equiv(k_\parallel,\vec k_\perp)$, where $k_\parallel$ is aligned with the
spectral axis $\nu$, and $\vec k_\perp$ lies in the plane of the sky.}
that may be occupied by smooth spectrum foreground emission.  At low-order $k_\parallel$ modes
within the limits of the wedge,
foreground contamination may be suppressed through a combination of calibration and model 
subtraction.  However, calibration or modeling errors rapidly re-establish the characteristic
wedge pattern.  Outside of the wedge, foreground contamination drops rapidly 
\citep{pober_et_al2013,thargarajan_et_al2015}, provided that the spectral responses
of antenna elements and analog electronics are sufficiently smooth\footnote{Here, we distinguish
the chromaticity of the elements in isolation from the chromaticity inherent to
element separation in an interferometric baseline.}.  In promising avoidance-based
foreground strategy employed by PAPER \citep{parsons_et_al2014,ali_et_al2015}, these modes
may be targeted as the lowest risk path for contraining 21~cm reionization in the near term.

In this paper, we examine the performance of a new prototype HERA element in the context of
this new wedge-based framework.  In coordination with \citet{XXXsisterpapers}, we compare 
reflectometry measurements with
a specification for the spectral performance of an element, given models of the relative brightness of
foreground and 21~cm reionization modes in $k$-space.  
% XXX document map
